		\begin{objectives}
			In this tutorial you work with \emph{span}, using its definition to justify
			your claims.

	These problems relate to the following Course Learning Objectives:
			\textit{Clearly and correctly express the mathematical ideas of linear algebra to others, and
			understand and apply logical arguments and definitions that have been written by others}.
		\end{objectives}

\subsection*{Problems}


\begin{enumerate}
	\item Consider the Predator-Prey (Lotka-Volterra) model with unknown parametres:
		$$
			F' = a \cdot R\cdot F - b \cdot F 
		$$
		$$
			R' = c \cdot R - d \cdot R \cdot F
		$$
	
	Recall that $a,b,c,d\geq 0$. 
	In class, we used the constants: $a=0.01, b=1.1, c=1.1, d=0.1$.
	
%	Equilibrium Solutions at:
%	\begin{itemize}
%		\item $(F,R)= (0,0)$
%		\item $(F,R) = \big(\frac{c}{d},\frac{b}{a}\big)$
%	\end{itemize}
	
	
	\begin{enumerate}
		\item From the phase plane graph, we can see that the solutions rotate clockwise. Explain with a real world rabbit vs foxes perspective why it makes sense to move clockwise and it wouldn't make sense to move counter-clockwise. \\
	
		\item Recall the real-world meaning of the parametres $a,b,c,d$ and recall the formulas for the equilibrium points for this model. 
		\item Find a different set of parametres $a,b,c,d$ that give rise to a different behaviour of solutions. 
		\item Explain the real-world implications/context of those parameters. \\

		\item From the phase plane, how can you identify the maximum number of foxes/rabbits? What about the minimum?
		\item Find a formula for the maximum and minimum number of foxes and rabbits. \\
		

		\item Using an implicit derivative, calculate $\frac{dF}{dR}$ as a function of $F$ and $R$. Solve the separable differential equation.
				$$
				-b \ln(y)+ay +dx-c\ln(x) = {\rm Constant}
				$$
		\item This means that the graphs of solutions in the phase plane, are level sets of a function $\Psi(x,y)$. What is this function $\Psi(x,y)$?
		\item Use this fact to find a formula for the max/min of the number of rabbits/foxes that depend only on the initial conditions $r_0$ and $f_0$ (and on the constants $a,b,c,d$).
		
		
	\end{enumerate}

	
	
		
\end{enumerate}